---
title: "R gmail"
output: html_notebook
---

#https://developers.google.com/gmail/api/quickstart/python

### configuracion
```{r}
library(gmailr)
```
#Client ID: 1083667676606-h13hhrb8bir2136pgk85c8ebrddts5f6.apps.googleusercontent.com
#Client Secret: 02gyDjl98sgI5AoK84AUmzgP

```{r}
gm_auth_configure(path = "C:/Users/ronal/Documents/credentials_gmail.json")
```

#correr esto en consola cuando se pide autorizacion por primera vez
```{r}
gm_auth()
```
#para reutilizar la autorizacion
```{r}
gm_auth(email = TRUE, cache = "C:/Users/ronal/.R/gargle/gargle-oauth")
```

### escribir mensajes

Create a new email with gm_mime() and the helper functions. When testing it is recommended to use gm_create_draft() to verify your email is formatted as you expect before automating it (if desired) with gm_send_message().
```{r}
test_email <-
  gm_mime() %>%
  gm_to("ronald.mego@comercio.com.pe") %>%
  gm_from("businessintelligencegec@gmail.com") %>%
  gm_subject("hola desde laptop local") %>%
  gm_text_body("Can you hear me now?")
```

#optional: You can add a file attachment to your message with gm_attach_file().
```{r}
#write.csv(mtcars,"mtcars.csv")
test_email <- test_email %>% gm_attach_file("mtcars.csv")
```

# Verify it looks correct
```{r}
gm_create_draft(test_email)
```
# Verify it looks correct
```{r}
gm_send_message(test_email)
```



### leer mensajes

# view the latest thread
```{r}
my_threads <- gm_threads(num_results = 10)
```

# retrieve the latest thread by retrieving the first ID
```{r}
latest_thread <- gm_thread(gm_id(my_threads)[[1]])
```

# The messages in the thread will now be in a list
```{r}
latest_thread$messages
```

# Retrieve parts of a specific message with the accessors
```{r}
my_msg <- latest_thread$messages[[1]]
```

```{r}
gm_to(my_msg)
gm_from(my_msg)
gm_date(my_msg)
gm_subject(my_msg)
#gm_body(my_msg)
```
# If a message has attachments, download them all locally with `gm_save_attachments()`.
```{r}
gm_save_attachments(my_msg,path='D:/Mego')
```




