
# https://michalbrys.gitbooks.io/r-google-analytics/content/chapter3/library_rga.html

# https://code.markedmondson.me/googleAnalyticsR/articles/management.html

# http://code.markedmondson.me/googleAnalyticsR/articles/setup.html#your-own-google-project
```{r}
library(googleAuthR)
library(googleAnalyticsR)
library(tidyverse)
```

```{r}
ga_auth(email='work@work.com') #correr en consola
```



## get a list of your accounts
```{r}
account_list <- ga_account_list() #- Get account summary including the ViewId
```
#get a data.frame of available dimensions and metrics, or just use meta directly to get the version that comes with the library.
```{r}
kpis<-ga_meta()
```

#listar custom variables
```{r}
ga_custom_vars_list(accountId='3055636', webPropertyId='UA-3055636-11')
```


#query demo de paginas vistas
```{r}
ga_pageviews <- google_analytics(viewId = 21928896, #vista
                                 date_range =c("2020-10-01","2020-10-24"), #rango de fechas
                                 metrics = "pageviews",  #metrica paginas vistas
                                 dimensions = "date", #dimension fecha
                                 max=-1) #sin limite
```




###OTRAS FUNCIONES UTILES###  para mas informacion mirar el link
```{r}
ga_account_list() #- Get account summary including the ViewId
ga_accounts() #- Get account metadata for your user
ga_webproperty() #- Get web property
ga_webproperty_list() #- List web properties for a particular accountId
ga_view() #- Get single View (Profile)
ga_view_list() #- List Views (Profile) for a particular accountId/webPropertyId
```

#custom variables
```{r}
ga_custom_vars_list(accountId='3055636', webPropertyId='UA-3055636-11') #- list all custom dimensions or metrics
ga_custom_vars() #- get meta data for a specific custom variable
ga_custom_vars_create() #- create a new custom variable
ga_custom_vars_patch() #- update an existing custom variable
```
#para especifiar los permisos requeridos
```{r}
#options(googleAuthR.scopes.selected = c("https://www.googleapis.com/auth/analytics",
#"https://www.googleapis.com/auth/drive"))
```


#listar reportes sin muestrear
```{r}
ga_unsampled_list(accountId='3055636', webPropertyId='UA-3055636-11', profileId='21928896') #profileid=viewid
```

#metadata del reporte sin muestrear
```{r}
ga_unsampled(accountId='3055636', webPropertyId='UA-3055636-11', profileId='21928896', unsampledReportId='-jVmHcZqQB2kInVInjlRAw')
```

#descarga del reporte sin muestrear (esto no funciona)
```{r}
df_premium_por_dia_gs<-ga_unsampled_download(reportTitle='Monetizables EC',accountId='3055636', webPropertyId='UA-3055636-11', profileId='21928896', downloadFile = FALSE)
```

