---
title: "R Notebook"
output: html_notebook
---

# https://anderfernandez.com/blog/reticulate-usar-python-en-r/
# https://stackoverflow.com/questions/52741484/how-to-import-pandas-using-r-studio

```{r}
library(reticulate)
```
instalaciones de python que tengo
```{r}
#conda_list()
```

```{r}
use_python("D:\\Temporal\\Anaconda\\python.exe",required = TRUE)
```


```{r}
py_config()
```
#para instalar paquetes de python
```{r}
#py_install("pandas")
```

```{python}
import numpy as np
```

```{python}
import pandas as pd
```

#creo la varialbe 'x' en python
```{python}
x=1
print(x)
```
# probando interactuar entre python y R

#llamo la variable 'x' desde R, le sumo 1 y lo grabo en 'y'
```{r}
z<-py$x+1
```

#llamo la variable 'y' desde Python y le sumo 1.
```{python}
w=r.z+1
print(w)
```

# week1
# The Python Programming Language: Functions

`add_numbers` is a function that takes two numbers and adds them together.

```{python}
x=1
y=2
x+y
```

```{python}
def add_numbers(x, y):
    return x + y

add_numbers(1, 2)
```
`add_numbers` updated to take an optional 3rd parameter. Using `print` allows printing of multiple expressions within a single cell.

```{python}
def add_numbers(x,y,z=None):
    if (z==None):
        return x+y
    else:
        return x+y+z

print(add_numbers(1, 2))
print(add_numbers(1, 2, 3))
```

`add_numbers` updated to take an optional flag parameter.
```{python}
def add_numbers(x, y, z=None, flag=False):
    if (flag):
        print('Flag is true!')
    if (z==None):
        return x + y
    else:
        return x + y + z
    
print(add_numbers(1, 2, flag=True))
print(add_numbers(1, 2, flag=False))
```
Assign function `add_numbers` to variable `a`.

```{python}
def add_numbers(x,y):
    return x+y

a = add_numbers
a(1,2)
```
# The Python Programming Language: Types and Sequences

Use `type` to return the object's type.
```{python}
type('This is a string')
type(None)
type(1)
type(1.0)
type(add_numbers)
```
Tuples are an immutable data structure (cannot be altered).
```{python}
x = (1, 'a', 2, 'b')
type(x)
```
Lists are a mutable data structure.

```{python}
x = [1, 'a', 2, 'b']
type(x)
```
Use `append` to append an object to a list.
```{python}
x.append(3.3)
print(x)
```
This is an example of how to loop through each item in the list.
```{python}
for item in x:
    print(item)
```
Or using the indexing operator:
```{python}
i=0
while( i != len(x) ):
    print(x[i])
    i = i + 1
```
Use `+` to concatenate lists.
```{python}
[1,2] + [3,4]
```
Use `*` to repeat lists.
```{python}
[1]*3
```
Use the `in` operator to check if something is inside a list.

```{python}
1 in [1, 2, 3]
```
Now let's look at strings. Use bracket notation to slice a string.

```{python}
x = 'This is a string'
print(x[0]) #first character
print(x[0:1]) #first character, but we have explicitly set the end character
print(x[0:2]) #first two characters
print(x[-1]) #This will return the last element of the string.
print(x[-4:-2]) #This will return the slice starting from the 4th element from the end and stopping before the 2nd element from the end.
print(x[:3]) #This is a slice from the beginning of the string and stopping before the 3rd element.
print(x[3:]) #And this is a slice starting from the 4th element of the string and going all the way to the end.
```

```{python}
firstname = 'Christopher'
lastname = 'Brooks'

print(firstname + ' ' + lastname)
print(firstname*3)
print('Chris' in firstname)
```
`split` returns a list of all the words in a string, or a list split on a specific character.
```{python}
firstname = 'Christopher Arthur Hansen Brooks'.split(' ')[0] # [0] selects the first element of the list
lastname = 'Christopher Arthur Hansen Brooks'.split(' ')[-1] # [-1] selects the last element of the list
print(firstname)
print(lastname)
```
Dictionaries associate keys with values.

```{python}
x = {'Christopher Brooks': 'brooksch@umich.edu', 'Bill Gates': 'billg@microsoft.com'}
x['Christopher Brooks'] # Retrieve a value by using the indexing operator
```

```{python}
x['Kevyn Collins-Thompson'] = None
x['Kevyn Collins-Thompson']
```

Iterate over all of the keys:
```{python}
for name in x:
    print(x[name])
```
Iterate over all of the values:
```{python}
for email in x.values():
    print(email)
```
Iterate over all of the items in the list:
```{python}
for name, email in x.items():
    print(name)
    print(email)
```
You can unpack a sequence into different variables:
```{python}
y = ('Christopher', 'Brooks', 'brooksch@umich.edu')
fname, lname, email = y

fname
lname
```

# The Python Programming Language: More on Strings
Python has a built in method for convenient string formatting.
```{python}
sales_record = { #dict
'price': 3.24,
'num_items': 4,
'person': 'Chris'}

sales_statement = '{} bought {} item(s) at a price of {} each for a total of {}'

print(sales_statement.format(sales_record['person'],
                             sales_record['num_items'],
                             sales_record['price'],
                             sales_record['num_items']*sales_record['price']))
```
`csv.Dictreader` has read in each row of our csv file as a dictionary. `len` shows that our list is comprised of 234 dictionaries.

```{python}
import csv

with open('D:/Google Drive/Miscursos/PYTHON/Coursera/course1_downloads/mpg.csv') as csvfile:
    mpg = list(csv.DictReader(csvfile))
   
len(mpg)
print()
mpg[:3] # The first three dictionaries in our list.

```
`keys` gives us the column names of our csv.
```{python}
mpg[0].keys()
```

This is how to find the average cty fuel economy across all cars. All values in the dictionaries are strings, so we need to convert to float.
```{python}
sum(float(d['cty']) for d in mpg) / len(mpg)
```
Similarly this is how to find the average hwy fuel economy across all cars.
```{python}
sum(float(d['hwy']) for d in mpg) / len(mpg)
```
Use `set` to return the unique values for the number of cylinders the cars in our dataset have.

```{python}
cylinders = set(d['cyl'] for d in mpg)
cylinders
```
Here's a more complex example where we are grouping the cars by number of cylinder, and finding the average cty mpg for each group.

```{python}
CtyMpgByCyl = []

for c in cylinders: # iterate over all the cylinder levels
    summpg = 0
    cyltypecount = 0
    for d in mpg: # iterate over all dictionaries
        if d['cyl'] == c: # if the cylinder level type matches,
            summpg += float(d['cty']) # add the cty mpg
            cyltypecount += 1 # increment the count
    CtyMpgByCyl.append((c, summpg / cyltypecount)) # append the tuple ('cylinder', 'avg mpg')

CtyMpgByCyl.sort(key=lambda x: x[0])
CtyMpgByCyl
```
Use `set` to return the unique values for the class types in our dataset.
```{python}
vehicleclass = set(d['class'] for d in mpg) # what are the class types
vehicleclass
```
And here's an example of how to find the average hwy mpg for each class of vehicle in our dataset.
```{python}
HwyMpgByClass = []

for t in vehicleclass: # iterate over all the vehicle classes
    summpg = 0
    vclasscount = 0
    for d in mpg: # iterate over all dictionaries
        if d['class'] == t: # if the cylinder amount type matches,
            summpg += float(d['hwy']) # add the hwy mpg
            vclasscount += 1 # increment the count
    HwyMpgByClass.append((t, summpg / vclasscount)) # append the tuple ('class', 'avg mpg')

HwyMpgByClass.sort(key=lambda x: x[1])
HwyMpgByClass
```
# The Python Programming Language: Dates and Times
```{python}
import datetime as dt
import time as tm
```


```{python}
tm.time()
```
Convert the timestamp to datetime.
```{python}
dtnow = dt.datetime.fromtimestamp(tm.time())
dtnow
```
Handy datetime attributes:
```{python}
dtnow.year, dtnow.month, dtnow.day, dtnow.hour, dtnow.minute, dtnow.second # get year, month, day, etc.from a datetime
```
timedelta` is a duration expressing the difference between two dates.
```{python}
delta = dt.timedelta(days = 100) # create a timedelta of 100 days
delta
```

`date.today` returns the current local date.
```{python}
today = dt.date.today()
```

```{python}
today - delta # the date 100 days ago
```

```{python}
today > today-delta # compare dates
```
# The Python Programming Language: Objects and map()

An example of a class in python:
```{python}
class Person:
    department = 'School of Information' #a class variable

    def set_name(self, new_name): #a method
        self.name = new_name
    def set_location(self, new_location):
        self.location = new_location
```


```{python}
person = Person()
person.set_name('Christopher Brooks')
person.set_location('Ann Arbor, MI, USA')
print('{} live in {} and works in the department {}'.format(person.name, person.location, person.department))
```
Here's an example of mapping the `min` function between two lists.
```{python}
store1 = [10.00, 11.00, 12.34, 2.34]
store2 = [9.00, 11.10, 12.34, 2.01]
cheapest = map(min, store1, store2)
cheapest
```
Now let's iterate through the map object to see the values.

```{python}
for item in cheapest:
    print(item)
```
#Here is a list of faculty teaching this MOOC. Can you write a function and apply it using map() to get 
#a list of all faculty titles and last names (e.g. ['Dr. Brooks', 'Dr. Collins-Thompson', …]) ?
```{python}
people = ['Dr. Christopher Brooks', 'Dr. Kevyn Collins-Thompson', 'Dr. VG Vinod Vydiswaran', 'Dr. Daniel Romero']
def split_title_and_name(person):
    title = person.split()[0]
    lastname = person.split()[-1]
    return '{} {}'.format(title,lastname)

list(map(split_title_and_name, people))
```

# The Python Programming Language: Lambda and List Comprehensions

Here's an example of lambda that takes in three parameters and adds the first two.

```{python}
my_function = lambda a, b, c : a + b
```

```{python}
my_function(1, 2, 3)
```

```{python}
#Convert this function into a lambda:
people = ['Dr. Christopher Brooks', 'Dr. Kevyn Collins-Thompson', 'Dr. VG Vinod Vydiswaran', 'Dr. Daniel Romero']

def split_title_and_name(person):
    return person.split()[0] + ' ' + person.split()[-1]

#option 1
for person in people:
    print(split_title_and_name(person) == (lambda x: x.split()[0] + ' ' + x.split()[-1])(person))

#option 2
list(map(split_title_and_name, people)) == list(map((lambda y: y.split()[0] + ' ' + y.split()[-1]),people))
```
Let's iterate from 0 to 99 and return the even numbers.

```{python}
my_list = []
for number in range(0, 100):
    if number % 2 == 0:
        my_list.append(number)
my_list
```

Now the same thing but with list comprehension.
```{python}
my_list = [number for number in range(0,100) if number % 2 == 0]
my_list
```

#Here, why don’t you try converting a function into a list comprehension.

```{python}
def times_tables():
    lst = []
    for i in range(10):
        for j in range (10):
            lst.append(i*j)
    return lst
times_tables()
```

```{python}
mytable = [i*j for i in range(10) for j in range(10)]
mytable
```

```{python}
times_tables() == mytable
```
# Here’s a harder question which brings a few things together.

# Many organizations have user ids which are constrained in some way. 
# Imagine you work at an internet service provider and the user ids are all two letters 
# followed by two numbers (e.g. aa49). Your task at such an organization might be 
# to hold a record on the billing activity for each possible user.

# Write an initialization line as a single list comprehension which creates a list of all possible user ids. 
# Assume the letters are all lower case.

```{python}
lowercase = 'abcdefghijklmnopqrstuvwxyz'
digits = '0123456789'

answer = [i+j+w+z for i in lowercase for j in lowercase for w in digits for z in digits]

answer[:10]
```

