---
title: "R Notebook"
output: html_notebook
---


```{r}
library(reticulate)
```
instalaciones de python que tengo
```{r}
#conda_list()
```

```{r}
use_python("D:\\Temporal\\Anaconda\\python.exe",required = TRUE)
```


```{r}
py_config()
```
#para instalar paquetes de python
```{r}
#py_install("pandas")
```

```{python}
import numpy as np
```

```{python}
import pandas as pd
```

## Creating Arrays
Create a list and convert it to a numpy array
```{python}
mylist = [1, 2, 3]
x = np.array(mylist)
x
```
Or just pass in a list directly
```{python}
y = np.array([4, 5, 6])
y
```
Pass in a list of lists to create a multidimensional array.
```{python}
m = np.array([[7, 8, 9], [10, 11, 12]])
m
```
Use the shape method to find the dimensions of the array. (rows, columns)
```{python}
m.shape
```
`arange` returns evenly spaced values within a given interval.
```{python}
n = np.arange(0, 30, 2) # start at 0 count up by 2, stop before 30
n
```
`reshape` returns an array with the same data with a new shape.
```{python}
n = n.reshape(3, 5) # reshape array to be 3x5
n
```

`linspace` returns evenly spaced numbers over a specified interval.
```{python}
o = np.linspace(0, 4, 9) # return 9 evenly spaced values from 0 to 4
o
```
`resize` changes the shape and size of array in-place.
```{python}
o.resize(3, 3)
o
```
`ones` returns a new array of given shape and type, filled with ones.
```{python}
np.ones((3, 2))
```
`zeros` returns a new array of given shape and type, filled with zeros.
```{python}
np.zeros((2, 3))
```
`eye` returns a 2-D array with ones on the diagonal and zeros elsewhere.

```{python}
np.eye(3)
```
`diag` extracts a diagonal or constructs a diagonal array.
```{python}
np.diag(y)
```
Create an array using repeating list (or see `np.tile`)
```{python}
np.array([1, 2, 3] * 3)
```
Repeat elements of an array using `repeat`.
```{python}
np.repeat([1, 2, 3], 3)
```
#### Combining Arrays

```{python}
p = np.ones([2, 3], int)
p
```
Use `vstack` to stack arrays in sequence vertically (row wise).
```{python}
np.vstack([p, 2*p])
```
Use `hstack` to stack arrays in sequence horizontally (column wise).
```{python}
np.hstack([p, 2*p])
```
## Operations
Use +, -, *, / and ** to perform element wise addition, subtraction, multiplication, division and power.

```{python}
print(x + y) # elementwise addition     [1 2 3] + [4 5 6] = [5  7  9]
print(x - y) # elementwise subtraction  [1 2 3] - [4 5 6] = [-3 -3 -3]
```

```{python}
print(x * y) # elementwise multiplication  [1 2 3] * [4 5 6] = [4  10  18]
print(x / y) # elementwise divison         [1 2 3] / [4 5 6] = [0.25  0.4  0.5]
```

```{python}
print(x**2) # elementwise power  [1 2 3] ^2 =  [1 4 9]
```
**Dot Product:**
[ð‘¥1 ð‘¥2 ð‘¥3]â‹…î€ˆî€†î€‡î€‡ð‘¦1ð‘¦2ð‘¦3î€‹î€‰î€Šî€Š=ð‘¥1ð‘¦1+ð‘¥2ð‘¦2+ð‘¥3ð‘¦3
```{python}
x.dot(y) # dot product  1*4 + 2*5 + 3*6
```

```{python}
z = np.array([y, y**2])
print(len(z)) # number of rows of array
```
Let's look at transposing arrays. Transposing permutes the dimensions of the array.
```{python}
z = np.array([y, y**2])
z
```
The shape of array `z` is `(2,3)` before transposing.
```{python}
z.shape
```
Use `.T` to get the transpose.
```{python}
z.T
```
The number of rows has swapped with the number of columns.
```{python}
z.T.shape
```
Use `.dtype` to see the data type of the elements in the array.
```{python}
z.dtype
```
Use `.astype` to cast to a specific type.
```{python}
z = z.astype('f')
z.dtype
```
## Math Functions
Numpy has many built in math functions that can be performed on arrays.
```{python}
a = np.array([-4, -2, 1, 3, 5])
```


```{python}
a.sum()
a.max()
a.min()
a.mean()
a.std()
a.argmax() #argmax` and `argmin` return the index of the maximum and minimum values in the array.
a.argmin()
```
## Indexing / Slicing
```{python}
s = np.arange(13)**2
s
```
Use bracket notation to get the value at a specific index. Remember that indexing starts at 0.

```{python}
s[0], s[4], s[-1]
```
Use `:` to indicate a range. `array[start:stop]`
Leaving start or stop empty will default to the beginning/end of the array.
```{python}
s[1:5]
```
Use negatives to count from the back.
```{python}
s[-4:]
```
A second `:` can be used to indicate step-size. `array[start:stop:stepsize]`
Here we are starting 5th element from the end, and counting backwards by 2 until the beginning of the array is reached.
```{python}
s[-5::-2]
```
Let's look at a multidimensional array.
```{python}
rt = np.arange(36)
rt.resize((6, 6))
rt
```

Use bracket notation to slice: `array[row, column]`
```{python}
rt[2,2]
```
And use : to select a range of rows or columns
```{python}
rt[3, 3:6]
```
Here we are selecting all the rows up to (and not including) row 2, and all the columns up to (and not including) the last column.
```{python}
rt[:2, :-1]
```
This is a slice of the last row, and only every other element.
```{python}
rt[-1, ::2]
```

We can also perform conditional indexing. Here we are selecting values from the array that are greater than 30. (Also see `np.where`)
```{python}
rt[rt > 30]
```

Here we are assigning all values in the array that are greater than 30 to the value of 30.
```{python}
rt[rt > 30] = 30
rt
```

## Copying Data
Be careful with copying and modifying arrays in NumPy!
r2 is a slice of rt

```{python}
r2 = rt[:3,:3]
r2
```
Set this slice's values to zero ([:] selects the entire array)
```{python}
r2[:] = 0
r2
```

`rt` has also been changed!
```{python}
rt
```
To avoid this, use `r.copy` to create a copy that will not affect the original array
```{python}
r_copy = rt.copy()
r_copy
```
Now when r_copy is modified, r will not be changed.

```{python}
r_copy[:] = 10
print(r_copy, '\n')
print(rt)
```
#test de lo aprendido
#cuando creas un array
```{python}
old = np.array([[1, 1, 1],
                [1, 1, 1]])
print(old)
```
# y lo igualas a otro nuevo array,y le cambias valores a dicho nuevo array
# tabmien se cambia en el array original
```{python}
new = old
new[0, :2] = 0
print(new)
print('\n')
print(old)
```
### Iterating Over Arrays
Let's create a new 4 by 3 array of random numbers 0-9.
```{python}
test = np.random.randint(0, 10, (4,3))
test
```
Iterate by row:

```{python}
for row in test:
    print(row)
```
Iterate by index:
```{python}
for i in range(len(test)):
    print(test[i])
```
Iterate by row and index:

```{python}
for i, row in enumerate(test):
    print('row', i, 'is', row)
```
Use `zip` to iterate over multiple iterables.
```{python}
test2 = test**2
test2
```

```{python}
for i, j in zip(test, test2):
    print(i,'+',j,'=',i+j)
```

```{python}
type(lambda x: x+1)
```

```{python}
(lambda x: x+1)(5)
```

```{python}
rt = np.arange(0, 36, 1)
rt
```


```{python}
rt=rt.reshape(6,6)
rt
```

```{python}
rt[::7]
```


```{python}
rt[:,::7]
```

```{python}
rt[0:6,::7]
```

```{python}
rt[2::2,2::2]
```

```{python}
rt[2:4,2:4]
```
